---
title: "Term Project"
author: "Chad Homan"
date: '2021-04-20'
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: bibliography.bib
---

```{r setup, echo=FALSE}
workdir <- system("git rev-parse --show-toplevel", intern=TRUE)
##workdir <- file.path(workdir, "completed/assignent04")
##workdir

## Set the working directory
##setwd(workdir)
##```{r setup}
knitr::opts_knit$set(root.dir = '/home/choman/repos/bellevue/dsc520/completed/assignment04')
```

# Abstract

You will be working on a research paper for your final project.  This project will include identifying a topic/problem that you want to solve using data science.  While the final solution to the problem does not need to be provided via programming – you will be doing some exploratory data analysis, transformations, and summary statistics on the data via R.  You are welcome to create a model based on what you have learned in this course to solve the problem, but this is not required.  Instead, a recommendation is required for a model or method you would implement to solve the problem.  There are 3 steps to this project, with the final deliverable being a formal paper (completed in R Markdown) that outlines the problem, shows the analysis done with the data, and concludes with your recommendation for next steps.  Each step provides detailed information that you must include in each phase of the project.

    Identify a topic or a problem that you want to research.  Provide an introduction that explains the problem statement or topic you are addressing. Why would someone be interested in this?  How is it a data science problem?
    Draft 5-10 Research questions that focus on the problem statement/topic.
    Provide a concise explanation of how you plan to address this problem statement.
    Discuss how your proposed approach will address (fully or partially) this problem.
    Do some digging and find at least 3 datasets that you can use to address the issue. (There is not a required number of fields or rows for these datasets)
        Original source where the data was obtained is cited and, if possible, hyperlinked.
        Source data is thoroughly explained (i.e. what was the original purpose of the data, when was it collected, how many variables did the original have, explain any peculiarities of the source data such as how missing values are recorded, or how data was imputed, etc.).
    Identify the packages that are needed for your project.

    What types of plots and tables will help you to illustrate the ﬁndings to your research questions?
    What do you not know how to do right now that you need to learn to answer your research questions?

#    Introduction

> Identify a topic or a problem that you want to research.  Provide an introduction 
> that explains the problem statement or topic you are addressing. Why would 
> someone be interested in this?  How is it a data science problem?


   
According to InvestmentU[@investmentu2021], the top 5 crypto currencies are:

- Bitcoin (BTC)
- Ethereum (ETH)
- Litecoin (LTC)
- Bitcoin Cash (BCH)
- Binance Coin (BNB)

This study will attempt to show a correlation with the precious metals market
Which will focus on:

- Palladium
- Platinum 
- Gold
- Silver

Since January, though we have seen the face of crypto market change. US News[@usnews2021] shows the top 7 crypto currencies to be:

- Bitcoin (BTC)
- Bitcoin Cash (BCH)
- Cosmos (ATOM)
- Dogecoin (DOG)
- Ethereum (ETH)
- Compound (COMP)
- Polkadot (DOT)


#    Research questions

- Is there a correlation with crypto currencies and precious metals?
- 
- 
- 
- 
- 

#    Approach

#    How your approach addresses (fully or partially) the problem.


#    Data  (Minimum of 3 Datasets - but no requirement on number of fields or rows)

The data used will be:

- Cryptocurrency Historical Prices https://www.kaggle.com/sudalairajkumar/cryptocurrencypricehistory
- Gold Prices (Kaggle) https://www.kaggle.com/tunguz/gold-prices
- Gold, Silver, S&P 500 Index & Barrick history https://www.kaggle.com/thelastwhitekid/gold-sp-500-index-arimax-time-series


#    Required Packages

```{r loadpackages, echo=FALSE, include=FALSE}
################################################################
#
# Start loading default operations
#
################################################################
## Load the ggplot2 package
# clean package loading based on 
# https://statisticsglobe.com/r-install-missing-packages-automatically
mypackages <- c("ggplot2", "pastecs", "plyr", "dplyr", "purrr", "stringr")
mypackages <- append(mypackages, c("readxl"))
mypackages <- append(mypackages, c("boot", "QuantPsyc"))
mypackages <- append(mypackages, c("relaimpo", "corrplot")) #, "Boruta"))
mypackages <- append(mypackages, c("car"))

not_installed <- mypackages[!(mypackages %in% installed.packages()[, "Package"])]
if(length(not_installed)) install.packages(not_installed)

# Load libraries
##lapply(mypackages, require, character.only=TRUE)
for (package in mypackages) {
  library(package, character.only=TRUE)
}

theme_set(theme_minimal())
binwidth <- .5
```

There are a number of packages used to help process this data. The list of packages 
are:

```{r show-packages, echo=FALSE}
mypackages
```


# Plots and Table Needs


# Questions for future steps

# References
