---
title: "Term Project Final: Step 3"
author: "Chad Homan"
date: '2021-06-02'
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: bibliography.bib
---

```{r setup, echo=FALSE}
since          <- "1960-01-01"
vote_count_min <- 2500
workdir <- system("git rev-parse --show-toplevel", intern=TRUE)
termdir <- file.path(workdir, "term")
datadir <- file.path(termdir, "data")
##workdir <- file.path(workdir, "completed/assignent04")
##workdir

## Set the working directory
##setwd(workdir)
###```{r setup}
knitr::opts_knit$set(base.dir = workdir) 
knitr::opts_knit$set(root.dir = workdir) 
knitr::opts_knit$set(term.dir = termdir) 
knitr::opts_knit$set(data.dir = datadir) ###'/home/choman/repos/bellevue/dsc520/completed/assignment04')
```

```{r loadpackages, echo=FALSE, include=FALSE}
################################################################
#
# Start loading default operations
#
################################################################
## Load the ggplot2 package
# clean package loading based on 
# https://statisticsglobe.com/r-install-missing-packages-automatically
mypackages <- c("ggplot2", "pastecs", "plyr", "dplyr", "purrr", "stringr")
mypackages <- append(mypackages, c("readxl"))
mypackages <- append(mypackages, c("boot", "QuantPsyc"))
mypackages <- append(mypackages, c("relaimpo", "corrplot")) #, "Boruta"))
mypackages <- append(mypackages, c("car", "hash"))

not_installed <- mypackages[!(mypackages %in% installed.packages()[, "Package"])]
if(length(not_installed)) install.packages(not_installed)

# Load libraries
##lapply(mypackages, require, character.only=TRUE)
for (package in mypackages) {
  library(package, character.only=TRUE)
}

theme_set(theme_minimal())
binwidth <- .5
```


# Introduction

Movies are a part of most peoples lives and many of us are into sequels and 
universes, sagas etc. We often know that a given movie sets all kinds of 
box office records. In fact, as of this report, Avatar[@wikiHighest] is the 
current highest grossing film world wide at $2,847,246,203. One can arguably say 
that as a result, Avatar is the best movie of all time. Although I would disagree
with that statement. When accounting for inflation, the same wikipedia page claims
Gone With The Wind is the best movie at $3,739,000,000 beating Avatar by 
$453,000,000 which came in at $3,286,000,000 (remember this is with inflation). 
Again I would disagree with this statement too, but I have never seen Gone With 
The Wind. Just because a movie grosses big at the box office does not 
necessarily mean it is the best movie.

However, in today's world with streaming services and subscription accounts, such
as Amazon and Netflix, it is becoming more difficult to say which is the best
top-rated, let alone the best overall "collection" of movies. Web sites 
like the Internet Movie Database (IMDB) and The Movie Database (TMDB) track movies
by popularity which arguably may be a better way to track the best movie of all 
time. In fact IMDB's the ultimate list says that The Godfather[@imdb100] is the best 
movie of all time, although it is a bit dated. TMDB ranks Cruella as the best
movie.

But what about collections? According to wikipedia, the Marvel Cinematic Universe
is the highest grossing franchise[@wikiCollections] at $8,525.3 million with 
Avengers: Endgame being the top single movie of that collection at $857.9 million.
Ironically, the Avengers franchise is the 4th highest franchise at $2,619.1 million
with, you guessed it, Avengers: Endgame being it's single most successful movie.
As a film going geek, I would agree with this. But again, this is based on gross
income and not popularity. For example, the most popular Netflix movie is We Can
Be Heroes[@netflix30], but no gross income. Hmmm!!


```{r loaddata, echo =FALSE}
myread.csv <-function(filename) {
  fullpath <- file.path(knitr::opts_knit$get()$data.dir, filename)
  tmpdf <- read.csv(fullpath)
  return(tmpdf)
}

tmdb_all <- myread.csv("tmdb_all_data.csv")
```


```{r data-cleaning, echo =FALSE, include=FALSE}
summary(tmdb_all)

print("Start cleaning tmdb all")

#head(tmdb_all$title)
#head(tmdb_all$belongs_to_collection)
#head(tmdb_all$release_date)
#head(tmdb_all$genres)
#head(tmdb_all$vote_average)
#head(tmdb_all$popularity)
#head(tmdb_all$runtime)
```


```{r dataclean1, echo =FALSE, include=FALSE}
names(tmdb_all)
```


```{r dataclean2, echo=FALSE, include=FALSE}
#tmdb_all$belongs_to_collection
nrow(tmdb_all)
```


```{r dataclean3, echo=FALSE, include=FALSE}
# 
# belongs_to_collection == 4
# genres == 6
# popularity == 13
# release_date == 17
# revenue == 18
# runtime == 19
# title == 23
# vote_average == 25
# vote_count == 26
tmdb_reduced <- tmdb_all[, c(4, 6, 13, 17, 18, 19, 23, 25, 26)]
```


```{r dataclean4, echo=FALSE, include=FALSE}
names(tmdb_reduced)
nrow(tmdb_reduced)
```


```{r dataclean5, echo=FALSE, include=FALSE}

# Reduced movie list by date: since
tmdb_reduced <- subset(tmdb_reduced, release_date >= as.Date(since))## 1000))
# Reduced movie list by vote_count: since
tmdb_reduced <- subset(tmdb_reduced, vote_count >= vote_count_min)

                       ##&& nchar(belongs_to_collection) > 0)

# remove things not in collections
tmdb_reduced <- tmdb_reduced %>% dplyr::filter(!(belongs_to_collection == ""))

```

```{r clean_results1, echo=FALSE, include=FALSE}
# Results of cleaning
dim(tmdb_all)
dim(tmdb_reduced)
```

```{r clean_results2, include=FALSE}
names(tmdb_all)
names(tmdb_reduced)
```

```{r clean_results3, echo=FALSE, include=FALSE}

head(tmdb_reduced)
head(tmdb_reduced$belongs_to_collection)
tmdb_reduced %>% dplyr::filter(grepl("Star Wars", belongs_to_collection))
tmdb_reduced %>% dplyr::filter(grepl("X-Men", belongs_to_collection))
tmdb_reduced %>% dplyr::filter(grepl("Marvel", belongs_to_collection))

#tmdb_reduced$belongs_to_collection[grepl("Star Wars Collection", tmdb_reduced$title)]

##rotten$year
##rotten$view_the_collection

##print ("Identified duplicates - based on 'type'")
##rotten$title[rotten$view_the_collection == "Marvel Cinematic Universe"]


##dimnames(rotten)

##head(rotten$critic_score)
##head(rotten$people_score)
##head(rotten$type)
##head(rotten$total_ratings)
##head(rotten$ratings)
##head(rotten$box_office_.gross_usa)
##head(rotten$runtime)


##marvel_universe <- rotten[rotten$view_the_collection == "Marvel Cinematic Universe",]
##dc_universe <- rotten[rotten$view_the_collection == "DC Extended Universe",]
##star_wars <- rotten[rotten$view_the_collection == "Star Wars Saga",]
##xmen <- rotten[rotten$view_the_collection == "X-men",]
##potter <- rotten[rotten$view_the_collection == "Harry Potter",]
##pixar <- rotten[rotten$view_the_collection == "Pixar",]


##marvel_universe$title
##star_wars$title
##potter$title
##pixar$title
##rotten$view_the_collection
##rotten[rotten$title == "Mission: Impossible"]


collections <- tmdb_reduced$belongs_to_collection %>% unique()
is.vector(collections)

mycollections = c()
for (i in 1:length(collections)) {
  x <- str_trim(strsplit(strsplit(collections[i], ",")[[1]][2], ":")[[1]][2])
  mycollections <- append(mycollections, x)
}

head(mycollections)
tmdb_reduced %>% dplyr::filter(grepl(mycollections[1], belongs_to_collection))
```

```{r build-somthing, echo=FALSE, include=FALSE}
print("BUILD SOMETHING")
head(tmdb_reduced)
length(mycollections)
head(mycollections)

## LOOP
for (i in mycollections) {
  print(i)
}

x <- data.frame()
for (i in mycollections) {
 # print(mycollections[i])
  tmp <- tmdb_reduced %>% dplyr::filter(grepl(i, belongs_to_collection))
  print("HEY YOU!!!!")
  print(tmp)
  #print(tmp)
  films <- nrow(tmp)
  votes_count.sum <- sum(tmp$vote_count)
  votes_average.ave <- sum(tmp$vote_average) / length(tmp)
  runtime.sum <- sum(tmp$runtime)
  runtime.ave <- sum(tmp$runtime) / length(tmp)
  #popularity.ave <- sum(tmp$popularity)/length(tmp)
  #popularity.ave <- sum(tmp$popularity)
  ##print(votes_count.sum)
  ##print(votes_average.ave)
  
  x <- rbind(x, data.frame(i, 
                votes_count.sum, votes_average.ave, 
                runtime.sum, runtime.ave, films))
}
#tmdb_reduced$popularity
colnames(x) <- c("collection", 
                 "vote_count_sum", "vote_count_ave", 
                 "runtime_sum", "runtime_ave", "count")

```

# The problem statement you addressed

So the problem statement is how to accurately determine the top 10 movies 
franchises by popularity, not gross as not all movies produce gross income.

# How you addressed this problem statement

Using the movie database from [https://www.kaggle.com/mathlasker/tmdballmovies](https://www.kaggle.com/mathlasker/tmdballmovies). This database currently contains `r prettyNum(nrow(tmdb_all), big.mark=",", scientific=FALSE)` movies. Movies are filtered down by:

- movies since `r since`
- vote counts above `r vote_count_min`
- by collections

Once I trimmed the database, I was able to build a database of just collection 
information that allows me to display the information in a way that is meaningful 
to me, a top ten list.

# Analysis / Implications
```{r top10, echo =FALSE}
# Gather top10 information
top10_vote  <- x[order(x$vote_count_sum, decreasing=TRUE), ]
top10_rt    <- x[order(x$runtime_sum, decreasing=TRUE), ]
top10_rtave <- x[order(x$runtime_ave, decreasing=TRUE), ]

# Gather top 2 information for report, seems like a function is needed
topmovie_vote        <- top10_vote$collection[1]
topmovie_vote_min    <- top10_vote$runtime_sum[1]
topmovie_vote_ave    <- round(top10_vote$runtime_ave[1])
topmovie_vote_count  <- top10_vote$count[1]
#
secondplace_vote     <- top10_vote$collection[2]
secondplace_vote_min <- top10_vote$runtime_sum[2]
secondplace_vote_ave <- round(top10_vote$runtime_ave[2])
secondplace_vote_count <- top10_vote$count[2]

beat_vote            <- top10_vote$vote_count_sum[1] - top10_vote$vote_count_sum[2]

# runtime
topmovie_rt           <- top10_rt$collection[1]
topmovie_rt_min       <- top10_rt$runtime_sum[1]
topmovie_rt_ave       <- top10_rt$runtime_ave[1]
topmovie_rt_count     <- top10_rt$count[1]
#
secondplace_rt        <- top10_rt$collection[2]
secondplace_rt_min    <- top10_rt$runtime_sum[2]
secondplace_rt_ave    <- top10_rt$runtime_ave[2]
secondplace_rt_count  <- top10_rt$count[2]

beat_rt_min           <- topmovie_rt_min - secondplace_rt_min

# runtime avergae
topmovie_ave        <- top10_rtave$collection[1]
topmovie_ave_min    <- top10_rtave$runtime_sum[1]
topmovie_ave_ave    <- top10_rtave$runtime_ave[1]
topmovie_ave_count  <- top10_rtave$count[1]
#
secondplace_ave        <- top10_rtave$collection[2]
secondplace_ave_min    <- top10_rtave$runtime_sum[2]
secondplace_ave_ave    <- top10_rtave$runtime_ave[2]
secondplace_ave_count  <- top10_rtave$count[2]
beat_ave_min           <- topmovie_ave_ave - secondplace_ave_ave
```

I was able to perform 3 different types of analysis on the data:

- Top 10 Movie Collection (by vote)
- Top 10 Movie Collection (by total runtime)
- Top 10 Movie Collection (by ruintime average)

Unfortunately I could not properly address some collections such as total 
universe collections (e.g. Marvel). The database need to be munged more and I 
lost time to do that. 

### Top 10 Movie Collections (by vote)

The analysis of the data shows that the `r topmovie_vote_count` movie 
`r topmovie_vote` with a total runtime of `r topmovie_vote_min` minutes 
(`r round(topmovie_vote_min / 60, digits=1)` hours), with an average of 
`r topmovie_vote_ave` minutes (`r round(topmovie_vote_ave / 60, digits=1)` hours)
per film beats the `r secondplace_vote_count` movie `r secondplace_vote` with a 
total collection runtime of `r secondplace_vote_min` minutes 
(`r round(secondplace_vote_min / 60, digits=1)` hours), with  an average of 
`r secondplace_vote_ave` minutes (`r round(secondplace_vote_ave / 60, digits=1)`
hours) per film by `r beat_vote` vote(s). 

*NOTE*: This however is straight movie collections and not total collections by 
universe (e.g. Marvel).

```{r top10display, echo =FALSE}
knitr::kable(head(top10_vote, n=10),
             col.names=c("Collection", 
                         "Vote Count", "Vote Ave", 
                         "Runtime Sum", "Runtime Ave", 
                         "Films"),
             row.names=FALSE,
             caption = "Top 10 Movies (By Vote)")

##x = head(top10_vote, n=5)
##ggplot(x, aes(x=collection, y=count)) + 
##  geom_bar(width=.5)

```

### Top 10 Movies Collections (by total runtime)

The analysis of the data shows that the `r topmovie_rt_count` movie 
`r topmovie_rt` with a total runtime of `r topmovie_rt_min` minutes 
(`r round(topmovie_rt_min / 60, digits=1)` hours), with an average of 
`r round(topmovie_rt_ave, digits=2)` minutes (`r round(topmovie_rt_ave / 60, digits=1)` hours) per film beats the `r secondplace_rt_count` movie `r secondplace_rt` with 
a total collection runtime of `r secondplace_rt_min` minutes 
(`r round(secondplace_rt_min / 60, digits=1)` hours), with  an average of 
`r round(secondplace_rt_ave, digits=2)` minutes 
(`r round(secondplace_rt_ave / 60,digits=1)` hours) per film by `r beat_rt_min`
minute(s). 

*NOTE*: This however is straight movie collections and not total collections by 
universe (e.g. Marvel).

```{r runtime_sum, echo=FALSE}
knitr::kable(head(top10_rt, n=10),
             col.names=c("Collection", 
                         "Vote Count", "Vote Ave", 
                         "Runtime Sum", "Runtime Ave", 
                         "Films"),
             row.names=FALSE,
             caption = "Top 10 Movies (Total Runtime)")
```

### Top 10 Movie Collections (runtime average)

The analysis of the data shows that the `r topmovie_ave_count` movie 
`r topmovie_ave` with a total runtime of `r topmovie_rt_min` minutes 
(`r round(topmovie_ave_min / 60, digits=1)` hours), with an average of 
`r round(topmovie_ave_ave, digits=2)` minutes (`r round(topmovie_ave_ave / 60, digits=1)` hours) per film beats the `r secondplace_ave_count` movie `r secondplace_ave` with a total collection runtime of `r secondplace_ave_min` minutes 
(`r round(secondplace_ave_min / 60, digits=1)` hours), with  an average of 
`r round(secondplace_ave_ave, digits=2)` minutes 
(`r round(secondplace_ave_ave / 60,digits=1)` hours) per film by 
`r round(beat_ave_min, digits=2)` minute(s). 

*NOTE*: This however is straight movie collections and not total collections by 
universe (e.g. Marvel).

```{r runtime_ave, echo =FALSE}
knitr::kable(head(top10_rtave, n=10),
             col.names=c("Collection", 
                         "Vote Count", "Vote Ave", 
                         "Runtime Sum", "Runtime Ave", 
                         "Films"),
             row.names=FALSE,
             caption="Top 10 Movies (Total Runtime Ave)")
```

# Limitations

- I do not feel I am addressing vote_count and vote_average correctly to 
  display the best movie franchise overall
- I am definitely not handling spinoffs correctly
- Personally I feel the one database is excellent, however there could be 
  additional columns
- Perhaps additional databases would be needed to see how the data skews.
- Adding gross incomes could help correlate the vote counts and averages
- There seems to be come broken data in the csv file that was noticed after 
  changing to the search parameter date of `r since`. 
- Since the collections are not all inclusive, it also makes determining 
  the most popular/best movie collection not accurate.

# Concluding Remarks

I forgot all about ktables(), I felt they produced a cleaner top ten chart. So I
am very happy with that. I also finally figured out how to embed R objects into
the text. This allows for easier generation of the PDF if the database is updated.

I feel that possibly bar charts may help explain the data better. Unfortunately,
I could not get them to compile. So I excluded them from my final product.

I think moving forward, more munging of the data is required to produce total 
collection and proper spinoff analysis. Also, the vote average needs to be 
understood better to produce a better ranking system. Adding additional databases
should help to improve the results by producing a more generic average ranking
system. Finally, Adding gross income and inflation adjusted income should help 
produce additional information. From this, additional graphs such as pie charts 
would help explain the story.

# References