---
title: "Term Project: Step 1"
author: "Chad Homan"
date: '2021-05-16'
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: bibliography.bib
---

```{r setup, echo=FALSE}
workdir <- system("git rev-parse --show-toplevel", intern=TRUE)
termdir <- file.path(workdir, "term")
datadir <- file.path(termdir, "data")
##workdir <- file.path(workdir, "completed/assignent04")
##workdir

## Set the working directory
##setwd(workdir)
###```{r setup}
knitr::opts_knit$set(base.dir = workdir) 
knitr::opts_knit$set(root.dir = workdir) 
knitr::opts_knit$set(term.dir = termdir) 
knitr::opts_knit$set(data.dir = datadir) ###'/home/choman/repos/bellevue/dsc520/completed/assignment04')
```

# Step 1

##    Introduction

Movies have been around for about a century or more. As a self-proclaimed movie
buff I am curious what currently are the most successful movie franchises? The
franchises will be evaluated as a whole and as a reboot. What do I mean by this?
For example, the Star Trek franchise as a whole would be all 13 Star Trek movies
ranked together. As a reboot, I would separate the 4 Picard Movies and the 3 
Newer Captain Kirk Movies from the original 6 Captain Kirk Movies. So another 
problem to solve is how many times was a franchise rebooted? And if 
possible with and without spin-offs. 

I believe many like-minded movie fans would be interested in this. Especially 
since there are many fan-duels between Star Wars vs Star Trek, Marvel vs DC, etc.

This is a data science problem because of the sheer amount of data and calculations
that will need to happen to solve the problem.

##    Research questions

> Draft 5-10 Research questions that focus on the problem statement/topic.

- How far back should I process data? 1960? 1970? or 1980 and above?
- How do I treat reboots?
- How do I identify spin-offs
- How do I address spin-offs
- How do I determine success of a movie? by ticket sales? by rating? by votes?
- How do I determine a movie that changes primary actor? as a reboot or continuance?
- Do I limit my rankings to the top 10 or 20?
- Will I need more the my original three datasets?
- Will I need to swap out any of my datasets?

### Aditional questions

- What are the most successful movie franchises? 
- What is the total run time of the movie franchises?
- How do the reboots affect success or not? 
- Do movie spin-offs have good success?
- Is the Marvel Universe stronger than the DC Universe in regards to success?
- How does the Star Wars franchise fair against the Star Trek franchise?
- How does the Star Trek franchises fair amongst themselves, old Kirk vs new Kirk vs Picard?
- How do the Star Wars spin offs fair by themselves?
- How does the Marvel franchise fair against the DC franchise (all together vs the universes)?
- How does Friday the 13th fair against Nightmare on Elm St? Should we look at other
  franchises like The Exorcist?
- Aliens vs Predator? vs ??
- James Bond vs Kingsman vs Mission Impossible?

##    Approach

1. Fist step is to gather the data (below) and possible other data sets not 
   mentioned.
2. Next is data scrubbing.
3. Start organizing the data into categories that will help answer the above 
   questions. 
4. Produce graphs and charts to visualize my solutions

##    How your approach addresses (fully or partially) the problem.

I believe this approach will partially address the problem. There will unforeseen
issues that will need to be addressed as I move forward with the solution.

## Data

Initially, four datasets have been identified to help solve this problem. They 
are:

- IMDB Movie Dataset
  - https://www.kaggle.com/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows
  - IMDB Dataset of top 1000 movies and tv shows
  - 16 Varibles
- Popular Movies of IMDB
  - https://www.kaggle.com/sankha1998/tmdb-top-10000-popular-movies-dataset
  - TMDb lists some 80,000 films at time of writing, which is considerably fewer than IMDb. While not as complete as IMDb, it holds extensive information for most popular/Hollywood films.
This is dataset of the 10,000 most popular movies across the world has been fetched through the read API.
  - 6 Variables
- IMDB Dataset
  - https://www.kaggle.com/manishkumarparmar/imdb-dataset
  - A Colection of movie information
  - 12 Varibles
- Rotten Tomatoes Top Movies
  - https://www.kaggle.com/fnunezsanchez/rotten-tomatoes-top-movies
  - Top-rated movies in different categories by critics and Rotten Tomatoes users
  - 26 Variables
- 

##    Required Packages

```{r loadpackages, echo=FALSE, include=FALSE}
################################################################
#
# Start loading default operations
#
################################################################
## Load the ggplot2 package
# clean package loading based on 
# https://statisticsglobe.com/r-install-missing-packages-automatically
mypackages <- c("ggplot2", "pastecs", "plyr", "dplyr", "purrr", "stringr")
mypackages <- append(mypackages, c("readxl"))
mypackages <- append(mypackages, c("boot", "QuantPsyc"))
mypackages <- append(mypackages, c("relaimpo", "corrplot")) #, "Boruta"))
mypackages <- append(mypackages, c("car"))

not_installed <- mypackages[!(mypackages %in% installed.packages()[, "Package"])]
if(length(not_installed)) install.packages(not_installed)

# Load libraries
##lapply(mypackages, require, character.only=TRUE)
for (package in mypackages) {
  library(package, character.only=TRUE)
}

theme_set(theme_minimal())
binwidth <- .5
```

There are a number of packages used to help process this data. The initial list 
of packages are:

```{r show-packages, echo=FALSE}
mypackages
```

## Plots and Table Needs

- histograms
- bar chart
- line graphs
- corrplots

## Questions for future steps

I have no question at the moment for future steps.


