---
title: "ASSIGNMENT 7.2"
author: "Chad Homan"
date: '2021-04-27'
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: bibliography.bib
---

```{r setup, echo=FALSE, include=FALSE}
mypackages <- c("ggplot2", "pastecs", "dplyr", "purrr", 
                "GGally", "reshape2", "scales", "knitr")

not_installed <- mypackages[!(mypackages %in% installed.packages()[, "Package"])]
if(length(not_installed)) install.packages(not_installed)

# Load libraries
##lapply(mypackages, require, character.only=TRUE)
for (package in mypackages) {
  library(package, character.only=TRUE)
}

workdir <- system("git rev-parse --show-toplevel", intern=TRUE)
knitr::opts_knit$set(root.dir = workdir) 
survey_df <- read.csv("student-survey.csv")

info <- function(df)
{
  print("**************************************************")
  print("Summary:")
  print(summary(df))
  print("")
  print("stat.desc:")
  print(stat.desc(df))
  print("**************************************************")
}

genderA <- survey_df[ which(survey_df$Gender == 0), ]
genderB <- survey_df[ which(survey_df$Gender == 1), ]
```

# Abstract

As a data science intern with newly learned knowledge in skills in statistical correlation and R programming, you will analyze the results of a survey recently given to college students. You learn that the research question being 
investigated is: “Is there a significant relationship between the amount of time spent reading and the time spent watching television?” You are also interested 
if there are other significant relationships that can be discovered? The survey 
data is located in this StudentSurvey.csv file.

This assignment will process the following table and attewmp to given correct
answers. 

```{r show-table, echo=FALSE}
kable(survey_df)
```

## Calculate Covariance

Use R to calculate the covariance of the Survey variables and provide an 
explanation of why you would use this calculation and what the results indicate.

*Answer*: Covariance is a statistical calculation that helps you understand how two sets of data are related to each other [@wiki2020]. In other words, it 
defines the changes between the two variables. The covariance of the survey data
is below:

```{r covAll}
cov(survey_df)
```

I would interpret this data to mean that there is a strong correlation between 
these variables. However, there is not a strong correlation when comparing any 
of the variables against gender. Also, there is a negative covariance when comparing TimeReading against TimeTV. There is also a negative covariance when comparing TimeReading against Happiness. However, there is a positive covariance 
when comparing TimeTV against Happiness.

If we process the genders:

```{r covGenders}
cov(genderA)
cov(genderB)
```

We notice that with the genders, the covariance expirenced above holds true per
gender. Although, genderA has a higher negative covariance than genderB when 
comparing TimeReading against TimeTV and TimeReading aginast Happiness. Oddly, genderA also has a higher covariance than genderB when comparing TimeTV against
Happiness.

## Survey Data Examination

1. Examine the Survey data variables.

There are four variables in the set of data

```{r variables, echo=FALSE}
colnames(survey_df)

print ("Statistic Info for: survey_df")
info(survey_df)

print ("Statistic Info for: genderA")
info(genderA)

print ("Statistic Info for: genderB")
info(genderB)

ggplot(data=survey_df, aes(x=TimeReading, y=TimeTV, group=1)) +
  geom_line(aes(y=TimeTV, color = "Both")) + 
  geom_line(data=genderA, aes(y=TimeTV, color = "GenderA")) +
  geom_line(data=genderB, aes(y=TimeTV, color = "GenderB")) +
  scale_color_manual("", 
       breaks = c("Both", "GenderA", "GenderB"),
       values = c("darkred", "darkgreen", "steelblue")) +
  xlab("Time Reading") + ylab("Time TV")
```

What measurement is being used for the variables? Explain what effect changing 
the measurement being used for the variables would have on the covariance calculation. Would this be a problem? Explain and provide a better alternative 
if needed.

## Corellation

Choose the type of correlation test to perform, explain why you chose this test,
and make a prediction if the test yields a positive or negative correlation?

### Perform a correlation analysis of:

#### All variables

* **Pearson Method**:

```{r correlation-p, echo =FALSE}
cor(survey_df)
pearson <- cor(survey_df)
GGally::ggpairs(survey_df)
```

* Kendall Method:

```{r correlation-k, echo =FALSE}
cor(survey_df, method ="kendall")
kendall <- cor(survey_df, method ="kendall")
```

* Spearman Method:

```{r correlation0s, echo =FALSE}
cor(survey_df, method ="spearman")
spearman <- cor(survey_df, method ="spearman")
```

* Spearman > Peason
```{r sp-compare, echo=FALSE}
spearman > pearson
```

Doing both is interesting because if you have Spearman > Pearson, that means 
that you have a correlation that is monotonic but not linear [@stack2010]. Therefore I believe pearson is a better option.

#### Single Correlation:

A single correlation between two a pair of the variables. The correlation between
TimeTV and TimeReading is:

```{r correlation2, echo =FALSE}
cor(survey_df$TimeTV, survey_df$TimeReading)
```

The correlation between TimeTV and Happiness is:

```{r correlation3, echo =FALSE}
cor(survey_df$TimeTV, survey_df$Happiness)
```


The correlation between TimeReading and Happiness is:

```{r correlation4, echo =FALSE}
cor(survey_df$TimeReading, survey_df$Happiness)
```

#### Confidence Interval


Repeat your correlation test in step 2 but set the confidence interval at 99%

```{r correlation5, echo =FALSE}
cor.test(survey_df$TimeTV, survey_df$TimeReading, conf.level = 0.99)
```

#### Correlation Matrix

Describe what the calculations in the correlation matrix suggest about the relationship between the variables. Be specific with your explanation.

```{r matrix, echo= FALSE}
cor(survey_df)
```

This show a strong negative correlation when comparing TimeReading with TimeTV, likewise a negative correlation when comparing TimeReading with Happiness. However, there is a strong positive correlation when comparing TimeTV with Happiness.

## correlation coefficient

Calculate the correlation coefficient and the coefficient of determination, describe what you conclude about the results.

Correlation coeficient on Time Reading vs Time TV is
```{r coef1, echo=FALSE}
r <- cor(survey_df$TimeReading, survey_df$TimeTV)
r
r^2
```

The coefficient of determination is:
```{r cod1, echo=FALSE}
cor.lm <- lm(TimeReading ~ TimeTV, data=survey_df)
summary(cor.lm)$r.squared
```

Personally, I am not too sure on what this means. However, according to Dan
Shiebler[@shiebler2017]

> The square of the correlation coefficient is generally equal to the coefficient of determination whenever there is no scaling or shifting of f that can improve the fit of f to the data. For this reason the differential between the square of the correlation coefficient and the coefficient of determination is a representation of how poorly scaled or improperly shifted the predictions f are with respect to y.

Since the coefficient shows a strongly negative score comparing TimeReading and TimeTV. Which should mean that these data points are moving apart. And that the coefficient squared equals the coefficient of determination. I would conclude that there is nothing that can help this data fit better.

## Analysis: TV vs Reading

Based on your analysis can you say that watching more TV caused students to read less? Explain.



## Partial Correlation

Pick three variables and perform a partial correlation, documenting which variable you are “controlling”. Explain how this changes your interpretation and explanation of the results.


# References

